---
title: "eds-212-day-2"
author: "Joe DeCesaro"
date: "8/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
```

### First order univariate derivatives

In your R Markdown document, find the derivative expression and slope at the indicated value for each the following functions in R (each in a separate code chunk).

A: $f(x)=x^2$ at x=3
```{r}
fx <- expression(x^2)

df_dx <- D(expr = fx, name = 'x')
df_dx

x <- 3
eval(expr = df_dx)
```

B: $P(a)=2(3a+2)^4-5$ at a=1.2
```{r}
Pa <- expression(2 * (3 * a + 2) ^4 - 5)

dP_da <- D(expr = Pa, name = 'a')
dP_da

a <- 1.2
eval(expr = dP_da)
```

C: $g(t)=(-4t)/(t^2+1)^3$ at t=0
```{r}
gt <- expression((-4 * t) / ((t ^2 + 1)^3))

dg_dt <- D(expr = gt, name = 't')
dg_dt

t <- 0
eval(expr = dg_dt)
```

### Higher order univariate derivatives
Back in RStudio, add a code chunk in your Rmd in which you find the second derivative with respect to z
for $G(z)=4.6−1.9z^3$
```{r}
Gz <- expression(4.6 - 1.9 * z^3)

dG_dz <- D(expr = Gz, name = "z")
dG_dz

d2G_dz2 <- D(expr = dG_dz, name = "z")

d2G_dz2
```

### Partial derivatives
Find all partials of the following function: $f(x,y,z)=2x^2y−3x+1+4z^4x$
```{r}
f_xyz <- expression((2 * x ^2 * y) - (3 * x) + 1 + (4 * z ^4 * x))

df_dx <- D(expr = f_xyz, name = "x")
df_dx

df_dy <- D(expr = f_xyz, name = "y")
df_dy

df_dz <- D(expr = f_xyz, name = "z")
df_dz
```

### Section 4: make some plots

```{r}
fx <- function(x) {
  
  2.4 * exp(-0.05 *x) * sin(x)
}

ggplot(data = data.frame(x = c(-0.5, 0.5)), aes(x = x)) + 
  stat_function(fun = fx)
```

```{r}
ggplot(data = penguins, aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point(aes(color = species)) +
  labs(x = "Bill length (mm)",
       y = "Bill depth (mm)",
       title = "Palmer Penguin Bill Measurements")

```

```{r}
ggplot(data = penguins, aes(x = species, y = flipper_length_mm)) +
  geom_jitter(aes(color = species)) +
  labs(x = "Species",
       y = "Flipper Length (mm)",
       title = "Palmer Penguin Flipper Measurements by Species")
```

